{\fontfamily{ptm}\selectfont The Proximal Point interpretation of Nesterov accelerated proximal gradient}

### 2024-09-18 Notebook LaTEX/Momentum Accelerations/paper.tex
>w<, sync

    \subsection{The line search algorithm idea}
        \par
        The Nesterov acceleration algorithm can not achieve linear convergence rate for function that has a strongly convexity constant $\mu > 0$. 
        This claim is showed in Aujol et.al \cite{aujol_optimal_2019}. 
        Variants of the Nesterov accelerated gradient method that can achieve linear rate of convergence exists but it requires knowledge of $\mu, L$ a priori. 
        \par
        Literatures attempts at improving the convergence of FISTA under strong convexity, and other characterizations are vast. 
        But it can be approximately categorized as two types: 
        \begin{enumerate}
            \item Heuristic regarding restarting the algorithm. 
            \item Line search and estaimating the parameters: $L, \mu$. 
        \end{enumerate}
        Beck and Toubolle \cite{beck_fast_2009} proposed the idea of line search for the Lipschitz constant and restarting the algorithm based on function value for assure monotone descending property. 
        Work by Su et.al \cite{su_differential_2015} proposed an ODE understanding and suggested provable linear convergence rate of Nesterov accelerated gradient algorithm under the presence of strong convexity. 
        The ODE understand provoked interest in the restart techniques of AG/FISTA and the approaches are diverse and fruitful.

@@

        \begin{algorithm}
            \caption{Routine 1}\label{alg:routine1}
            \begin{algorithmic}[1]
                \Procedure{Routine1}{$L, h=f+g, x_0 = y_0 = z_0, \epsilon > 0$} %\Comment{The gcd of $a$ and $b$}
                    \State \textbf{Initialize} $\tilde \mu = 0, \eta_0 = 0, \tilde \eta_1 = L^{-1}$
                    \For{$t = 1, \cdots, N$}
                        \State $\tilde x \gets (\tilde \mu\tilde \eta_{t + 1} + 1)^{-1}(\tilde \mu \tilde \eta_{t + 1}y_t + x_t - \tilde \eta_{t + 1}\mathcal G_l(y_t))$
                        \If{$\exists \tilde \mu \in [\epsilon, L]: l_h(\tilde x, x_t, \tilde \mu) \le h(\tilde x)$}
                            \State \textbf{Find } $\tilde \mu \in [\epsilon, L]$ s.t: $l_h(\tilde x, x_t, \tilde \mu) \le h(\tilde x)$ \Comment{Use binary search}
                            \State $\tilde \eta_{t + 1} \gets (\sqrt{L\tilde \mu} - \tilde \mu)^{-1}$
                        \Else \Comment{$\tilde \mu = 0$}
                            \State $\tilde \mu \gets 0$
                            \State $\tilde \eta_{t + 1} \gets \theta_tL^{-1}$
                        \EndIf
                        \State $\eta_t \gets \tilde \eta_{t + 1} - L^{-1} - L^{-1}\mu\tilde\eta_{t + 1}$
                        \State $z_{t + 1} \gets y_t - L^{-1}\mathcal G_L(y_t)$
                        \State $x_{t + 1} \gets (\tilde \mu\tilde \eta_{t + 1} + 1)^{-1}(\tilde \mu \tilde \eta_{t + 1}y_t + x_t - \tilde \eta_{t + 1}\mathcal G_l(y_t))$
                        \State $y_{t + 1} \gets (1 + L\eta_{t + 1})^{-1}(L\eta_{t + 1}z_{t + 1} + x_{t + 1})$
                    \EndFor
                \EndProcedure
            \end{algorithmic}
        \end{algorithm}
        The quantity $\epsilon$ is the tolerance for line search to terminate if it can't find any such value.

>>>>

    \subsection{The line search algorithm idea}
        \par
        The Nesterov acceleration algorithm can not achieve linear convergence rate for function that has a strongly convexity constant $\mu > 0$. 
        This claim is showed in Aujol et.al \cite{aujol_optimal_2019}. 
        Variants of the Nesterov accelerated gradient method that can achieve linear rate of convergence exists but it requires knowledge of $\mu, L$ a priori. 
        \par
        Literatures attempts at improving the convergence of FISTA under strong convexity, and other characterizations are vast. 
        But it can be approximately categorized as two types: 
        \begin{enumerate}
            \item Heuristic regarding restarting the algorithm. 
            \item Line search and estaimating the parameters: $L, \mu$. 
        \end{enumerate}
        Beck and Toubolle \cite{beck_fast_2009} proposed the idea of line search for the Lipschitz constant and restarting the algorithm based on function value for assure monotone descending property. 
        Work by Su et.al \cite{su_differential_2015} proposed an ODE understanding and suggested provable linear convergence rate of Nesterov accelerated gradient algorithm under the presence of strong convexity. 
        The ODE understandings provoked interest in the restart techniques of AG/FISTA and the approaches are diverse and fruitful. 
        For example Aujol et.al \cite{aujol_parameter-free_2023,aujol_fista_2022}, Luca et.al \cite{calatroni_backtracking_2019} proposed methods of restarting the FISTA algorithm with an estimate to the strong convexity constant. 
        They proved the convergence rate in $\bar L$ and $\rho$ which are the estimates of Lipschitz constant $L$ through out the algorithm and $\rho \in (0, 1)$ another parameter that assist with the search for strong convexity constant $\mu$.

@@

        \begin{algorithm}
            \caption{Routine 1}\label{alg:routine1}
            \begin{algorithmic}[1]
                \Procedure{Routine1}{$L, \tilde \mu,h=f+g, (x_i, y_i), \epsilon > 0, N\in N, i \in \N$}
                    \State \textbf{Initialize: } $\eta_i = 0, \tilde \eta_{i + 1} = \eta_i + L^{-1} + \tilde \mu L^{-1} \eta_t$
                    \State $\tilde x \gets (\tilde \mu\tilde \eta_{i + 1} + 1)^{-1}(\tilde \mu \tilde \eta_{i + 1}y_i + x_i - \tilde \eta_{i + 1}\mathcal G_l(y_t))$
                    \If{$\exists \tilde \mu \in [\epsilon, L]: l_h(\tilde x, x_t, \tilde \mu) \le h(\tilde x)$ }
                        \State \textbf{find } $\tilde \mu \in [\epsilon, L]$ s.t: $l_h(\tilde x, x_t, \tilde \mu) \le h(\tilde x)$  
                        \State \textbf{updates} $\tilde \mu$ accordingly. 
                        \Comment{Use binary search}
                        \For{$j = i, \cdots, N$} 
                            \State $\tilde \eta_{t + j} \gets (\sqrt{L\tilde \mu} - \tilde \mu)^{-1}$
                        \EndFor
                    \Else \Comment{$\tilde \mu = 0$}
                        \State $\tilde \mu \gets 0$
                        \For{$j = i, \cdots, N$}
                            \State $\tilde \eta_{t + i} \gets \theta_{j}L^{-1}$
                        \EndFor
                    \EndIf
                    \For{$t = i, \cdots, N$}
                        \State $\eta_t \gets \tilde \eta_{t + 1} - L^{-1} - L^{-1}\mu\tilde\eta_{t + 1}$
                        \State $z_{t + 1} \gets y_t - L^{-1}\mathcal G_L(y_t)$
                        \State $x_{t + 1} \gets (\tilde \mu\tilde \eta_{t + 1} + 1)^{-1}(\tilde \mu \tilde \eta_{t + 1}y_t + x_t - \tilde \eta_{t + 1}\mathcal G_l(y_t))$
                        \State $y_{t + 1} \gets (1 + L\eta_{t + 1})^{-1}(L\eta_{t + 1}z_{t + 1} + x_{t + 1})$
                    \EndFor
                \EndProcedure
            \end{algorithmic}
        \end{algorithm}
        In above 
        \hyperref[alg:routine1]
            {Routine \ref*{alg:routine1}}
        we have 
        \begin{enumerate}
            \item $\epsilon$ is the tolerance for line search to terminate if it can't find any such value. 
            \item $(x_i, y_i)$ be any two iterates. 
            \item $\tilde \mu$ is any estimate of the strong convexity constant. It should be less than $L$ which is the true Lipschitz smoothness. 
            \item $i\in \N$ is an index that indicates $(x_i, y_i)$ could already be generated by Routine 1 in prior calls to the Routine 1. 
        \end{enumerate}
        Routine 1 estimates $\mu$ then performs accelerated gradient, the strong convexity index locally it then determine the stepsize for $\tilde \eta_t, \eta_t$ to fit either the traditional FISTA algorithm, or their strong convexity variants. 
        It executes for a fixed number of steps $N$ and it repeats periodically to check and update the strong convexity parameter. 
        However, there are many alternative ways of determining the strong convexity index $\mu$ while performing AG. 
        When evaluating function value is costly, alternatves that doesn't involve the use of function will accelerates the algorithm. 
        We propose the following method of estimating the strong convexity constant without using the function value. 
        Consider any fixed $x, y \in \RR^n$, then 
        \begin{align*}
            \tilde \mu = 
            \frac{
                \langle \mathcal G_L(x) - \mathcal G_L(y), x - y\rangle
            }{\Vert x - y\Vert^2}
        \end{align*}
        provides an estimate.
